---
title: "DECEA"
author: "RQ"
date: "07/07/2020"
output: html_document
---


```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(trrrj)
```

downloaded trajectory data for SBSP and stored locally, i.e. 2019-09-01 - 2019-09-05.
bounding box defined to cover radius of 102NM

define ARP for sbsp

```{r}
arp_sbsp <- data.frame(icao = "SBSP", lat = -23.626110077, lon= -46.65638733, elev = 2631)
```

read in the downloaded data

```{r}
pth <- "./data-raw/SBXX/"
fns <- list.files(path = pth, pattern = "^osn_SBSP_")

osn_col_spec <- cols(
  time = col_double(),
  icao24 = col_character(),
  lat = col_double(),
  lon = col_double(),
  velocity = col_double(),
  heading = col_double(),
  vertrate = col_double(),
  callsign = col_character(),
  onground = col_logical(),
  alert = col_logical(),
  spi = col_logical(),
  squawk = col_double(),
  baroaltitude = col_double(),
  geoaltitude = col_double(),
  lastposupdate = col_double(),
  lastcontact = col_double(),
  hour = col_double()
)

ds  <- paste0(pth, fns) %>%
  purrr::map_dfr(.f=~read_csv(., col_types = osn_col_spec))

ds <- ds %>% mutate(time = lubridate::as_datetime(time))
```


```{r}
pick <- c("e48e78","e4926f","e4827c","e49270")

picks <- ds %>% filter(time <= lubridate::ymd_hms("2019-09-01 15:00:00")
                       ,icao24 %in% pick)
```


```{r}
ggplot() +
  geom_point(data = picks, mapping = aes(x = lon, y = lat, colour = icao24))
```

```{r}
ggplot() +
  geom_point(data = picks, mapping = aes(x = time, y = geoaltitude, colour = icao24))
```


for testing purposes restrict to before noon = all arrivals

```{r}
picks_arr <- picks %>% filter(time <= lubridate::ymd_hms("2019-09-01 12:00:00")) %>%
  group_by(icao24) %>%
  arrange(time, .by_group = TRUE) %>%
  mutate(dist_arp = geosphere::distHaversine(
                        p1 = c(arp_sbsp$lon, arp_sbsp$lat)
                       ,p2 = cbind(lon, lat))
      , arp_brg = geosphere::bearing(
                        p1 = c(arp_sbsp$lon, arp_sbsp$lat)
                       ,p2 = cbind(lon, lat))
      ) %>%
  # convert to NM
  mutate( dist_arp = dist_arp / 1852)

asma_df <- picks_arr %>% 
    select(icao24, callsign, time, dist_arp, arp_brg) %>% 
    filter(dist_arp >= 99.8 & dist_arp <= 100.2) %>%
    summarise(time = median(time), dist_arp = median(dist_arp), brg = median(arp_brg) )

asma_df
```

```{r}
ggplot() +
  geom_histogram(data = asma_df, mapping = aes(x = brg), binwidth = 30)
```











develop quick heuristic to check for arrivals at SBSP

```{r}
near_apt <- function(.trjs, .near_alt = 750, .near_arp = 8000){
  candidates <- .trjs %>% 
    group_by(icao24, callsign) %>%
    summarise(
       near_min_alt = min(geoaltitude, na.rm = TRUE)
      ,near_min_darp= min(dist_arp, na.rm = TRUE)
      ) %>%
    filter(near_min_alt <= .near_alt, near_min_darp <= .near_arp) %>%
    ungroup()
  return(candidates)
}
```

check if flights are near SBSP, i.e. within 15NM

sbsp_15NM <- polygon_at_distance(c(arp_sbsp$lon, arp_sbsp$lat), 15) 
sbsp_15NM <- sbsp_15NM %>% st_sf %>% mutate(name = "SBSP 15NM")

convert positions to sf

pos <- ds %>% select(icao24, lat, lon) %>%
   st_as_sf(coords = c("lon","lat"), crs = 4326, agr = "constant")

cands <- st_join(pos, sbsp_15NM, join = st_within)




==========================

rq <- ds %>% filter(time <= lubridate::ymd_hms("2019-09-02 24:00:00", tz = "UTC")) %>%
  group_by(icao24)%>%
  arrange(time, .by_group = TRUE) %>%
  mutate(dt = time - lag(time)
        ,dist = geosphere::distHaversine(p1 = cbind(lon, lat), cbind(lag(lon), lag(lat)))) %>% 
  select(time, icao24, lat, lon, dt, dist, everything() ) %>%
  ungroup()

## basic prep
rq <- rq %>% 
  mutate(  dt = ifelse(is.na(dt), 0, dt)
        ,dist = if_else(is.na(dist), 0, dist)
        , leg = if_else(dt == 0 | dt >= 300, 1, 0)) %>%
  select(time, icao24, leg, lat, lon, dt, dist, everything() )

rq <- rq %>% group_by(icao24) %>% arrange(time, .by_group = TRUE) %>%
  mutate(leg = cumsum(leg)) %>%
  ungroup()

flts_sf <- rq %>% group_by(icao24, leg) %>% arrange(time, .by_group = TRUE) %>%
  select(icao24, leg, dt, dist, lat, lon, callsign, onground) %>% ungroup
  
flts_sf <- st_as_sf(flts_sf, coords = c("lon","lat"), crs = 4326, agr = "constant")

--> check does not look like a proper sf object
  
sbsp_15NM <- polygon_at_distance(c(arp_sbsp$lon, arp_sbsp$lat), 15) 
sbsp_15NM <- sbsp_15NM %>% st_sf %>% mutate(name = "SBSP 15NM")  

cands <- st_join(flts_sf, sbsp_15NM, join = st_within)

cands2 <- cands %>% filter(!is.na(name)) %>% group_by(icao24, leg)
