---
title: "VTBS"
author: "Performance Review Unit"
date: "02/06/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
pth <- "./data-raw/"
fns <- list.files(path = pth, pattern = "^osn_VTBS_")

osn_col_spec <- cols(
  time = col_double(),
  icao24 = col_character(),
  lat = col_double(),
  lon = col_double(),
  velocity = col_double(),
  heading = col_double(),
  vertrate = col_double(),
  callsign = col_character(),
  onground = col_logical(),
  alert = col_logical(),
  spi = col_logical(),
  squawk = col_double(),
  baroaltitude = col_double(),
  geoaltitude = col_double(),
  lastposupdate = col_double(),
  lastcontact = col_double(),
  hour = col_double()
)

ds  <- paste0(pth, fns) %>%
  purrr::map_dfr(.f=~read_csv(., col_types = osn_col_spec))

ds <- ds %>% mutate(time = lubridate::as_datetime(time))
```


 ICAO    LON   LAT  ELEV
  <chr> <dbl> <dbl> <dbl>
1 VTBS   101.  13.7     5

https://skyvector.com/airport/VTBS/Suvarnabhumi-International-Airport
converted with: https://www.rapidtables.com/convert/number/degrees-minutes-seconds-to-degrees.html

```{r}
arp_vtbs <- data.frame(icao = "VTBS", lat = 13.68583, lon = 100.7488)
arp_vtbd <- data.frame(icao = "VTBD", lat = 13.912599, lon= 100.6070)
```

```{r}
tmp <- ds %>% filter(geoaltitude <= 1000)

tmp2<- tmp %>% 
  group_by(icao24, callsign) %>% 
  summarise( N = n()
            ,min_time = min(time, na.rm = TRUE), max_time = max(time, na.rm = TRUE)
            ,ong = sum(onground, na.rm = TRUE))
```
  
```{r}
ggplot(data = tmp, mapping = aes(x = lon, y = lat)) + 
  geom_point() +
  geom_point(data = arp_vtbs, mapping = aes(x = lon, y = lat), colour = "green", size = 6)
```

```{r}
picks <- c("880446","3c15be","7813b5","8991b8","7582e3")
flts  <- ds %>% 
  filter(icao24 %in% picks)

ggplot() + 
  geom_point(
     data    = flts
    ,mapping = aes(x = lon, y = lat, colour = callsign)
  ) +
  geom_point(
     data    = arp_vtbs
    ,mapping = aes(x = lon, y = lat), colour = "green", size = 6
    ) +
  geom_point(
     data    = arp_vtbd
    ,mapping = aes(x = lon, y = lat), colour = "purple", size = 6
    )
```

```{r}
ggplot() +
  geom_point(
     data    = flts
    ,mapping = aes(x = time, y = geoaltitude, colour = callsign)
  )
```

```{r}
dlh792_793 <- flts %>% filter(icao24 == "3c15be")

append_dist_arp <- function(.trjs, .arp_lat, .arp_lon){
  df <- .trjs %>%  
    mutate(
       arp_lat  = .arp_lat , arp_lon = .arp_lon
      ,dist_arp = geosphere::distHaversine( p1 = cbind(lon, lat)
                                           ,p2 = cbind(arp_lon,arp_lat)) 
      )
  return(df)
}  
 
dlh792_793 <- dlh792_793 %>% append_dist_arp(arp_vtbs$lat, arp_vtbs$lon)

ggplot() +
  geom_point(
     data    = dlh792_793
    ,mapping = aes(x = time, y = geoaltitude, colour = callsign)
  )
```

```{r}
ggplot() +
  geom_point(
    data = dlh792_793
    ,mapping = aes(x = time, y = dist_arp, colour = callsign)
  )
```

distance-altitude

```{r}
g <- ggplot() +
  geom_point(
    data = dlh792_793
    ,mapping = aes(x = dist_arp,  y = geoaltitude, colour = callsign)
  ) +
  geom_vline(xintercept = 10000, colour = "green") +
  geom_hline(yintercept = 800,  colour = "green") + 
  theme_minimal()
g
```

zoom

```{r}
g + xlim(0, 15000) + ylim(0, 1500)
```

downsample for arrival and departure candidates

* filter trajectories for alt below threshold and distance arp below threshold

```{r}
near_apt <- function(.trjs, .near_alt = 750, .near_arp = 8000){
  candidates <- .trjs %>% 
    group_by(icao24, callsign) %>%
    summarise(
       near_min_alt = min(geoaltitude, na.rm = TRUE)
      ,near_min_darp= min(dist_arp, na.rm = TRUE)
      ) %>%
    filter(near_min_alt <= .near_alt, near_min_darp <= .near_arp) %>%
    ungroup()
  return(candidates)
}
```

```{r}
dlh792_793 %>% near_apt()
```

```{r}
vtbs_tfc <- ds %>% 
  append_dist_arp(arp_vtbs$lat, arp_vtbs$lon) %>%
  near_apt()
```

A tibble: 1,176 x 4
icao24 callsign near_min_alt near_min_darp
   <chr>  <chr>           <dbl>         <dbl>
 1 040045 ETH608           236.         4804.
 2 040045 ETH609           244.         5188.
 3 040045 NA               343.         4717.
 4 04c11a KQA886           190.         5425.
 5 04c11a KQA887           236.         6259.
 6 04c11a NA               297.         5450.
 
results show tibble with various NAs --> need to clean trajectory and fill blanks
on callsign

omitting NA
> vtbs_tfc %>% na.omit()
# A tibble: 926 x 4
   icao24 callsign near_min_alt near_min_darp
   <chr>  <chr>           <dbl>         <dbl>
 1 040045 ETH608           236.         4804.
 2 040045 ETH609           244.         5188.
 3 04c11a KQA886           190.         5425.
 4 04c11a KQA887           236.         6259.
 5 05a054 DQA364           267.         4782.
 6 05a05d DQA378           251.         5504.
 
 do some cleaning with 040045 ETH608 and 609
 
```{r}
eth608_609 <- ds %>% filter(icao24 == "040045")
colSums(is.na(eth608_609))
```

"only" 4 callsigns are missing!
 
ggplot() + geom_point(data = ds %>% filter(icao24 %in% c("899121","884341")), mapping = aes(x = lon, y = lat, colour = icao24)) + geom_point(data = arp_vtbs, mapping = aes(x = lon, y = lat), colour = "green", size = 5)

```{r}
append_dtime <- function(.trjs){
  trjs <- .trjs %>% 
    group_by(icao24) %>%
    arrange(time) %>%
    mutate(
      dtime = time - lag(time)
     ,dtime = if_else(is.na(dtime), as.difftime(0, units = "secs"), dtime)  
     ) %>%
    ungroup()
}

append_leg <- function(.trjs, .max_leg_gap = 300){
  trjs <- .trjs %>%
    group_by(icao24) %>% arrange(time) %>%
    mutate(
     leg = as.numeric(dtime) > .max_leg_gap | as.numeric(dtime) == 0
    ,leg = cumsum(leg == TRUE)
    ) %>%
    ungroup()
}

fill_callsign <- function(.trjs){
  trjs <- .trjs %>%
    group_by(icao24) %>% arrange(time) %>%
    tidyr::fill(callsign, .direction = "updown") %>%
    ungroup()
}
```

```{r}
tfc <- ds %>%
  filter(icao24 %in% vtbs_tfc$icao24) %>%
  append_dist_arp(arp_vtbs$lat, arp_vtbs$lon) %>%
  append_dtime() %>%
  append_leg() %>%
  fill_callsign()
```

do some stats

```{r}
tfc_stats <- tfc %>% group_by(icao24, callsign) %>%
  summarise(
     hits = n()
    ,legs = length(unique(leg))
  )
```

check direction of flight: going to / from airport, never are closer than x


GARBAGE - NOT SURE WHY - KEEP FOR A WHILE =============================

DONE - break down into 2 functions
- identify leg(s)
- fill callsign

bkp <- ds %>% filter(icao24 == "884341") %>%
  append_dtime() %>%
  mutate(
     leg = as.numeric(dtime) > 300 | as.numeric(dtime) == 0
    ,leg = cumsum(leg == TRUE)
    ) %>%
  group_by(icao24, leg) %>%
  tidyr::fill(callsign, .direction = "downup") %>%
  ungroup()

bkp <- ds %>% filter(icao24 == "884341") %>%
  filter(icao24 %in% vtbs_tfc$icao24) %>%
  append_dtime() %>%
  append_leg() %>%
  fill_callsign()

